Non-kissing and non-crossing complexes \\ for locally gentle algebras

### 2018-07-12 draft/accordionComplexSurfaces.tex
surface

\begin{remark}
\label{rem:propertiesSurface}
The following observations are useful for the computation of examples.
\begin{enumerate}[(i)]
\item The set~$V_{\bar Q}$ has one vertex for each straight walk in~$\bar Q$ (equivalently, for each maximal path in~$\bar Q$).
      Finite straight walks yield vertices on the boundary of~$\surface_{\bar Q}$, while infinite cyclic straight walks in~$\bar Q$ yield punctures~$\surface_{\bar Q}$ in~$V_{\bar Q}$.
      We denote by~$p$ the number of infinite cyclic straight walks in~$\bar Q$.
\item \label{item:arcs}
      The dissection~$\dissection_{\bar Q}$ has one edge for each vertex~$v \in \bar Q_0$, obtained by concatenation of the sides~$\Enrt{\alpha} = \Enrs{\beta}$ and~$\Enrt{\alpha'} = \Enrs{\beta'}$ where~$v = t(\alpha) = s(\beta) = t(\alpha') = s(\beta')$, $\alpha\beta \notin I$ and~$\alpha'\beta' \notin I$. We denote by~$\edgeof(v)$ the edge of~$\dissection_{\bar Q}$ corresponding to~$v$.
\item The dissection~$\dissection_{\bar Q}$ has one $\ell$-cell for each straight walk of length~$\ell$ in~$\bar Q\koszul$.
\item Similar statements hold dually for~${{V\dual}\!\!_{\bar Q}}$ and~${\dissection\dual\!\!_{\bar Q}}$.
\item The number of punctures of~$\surface_{\bar Q}$ is the number~$p + p\dual$ of infinite straight walks in~$\bar Q$ and in~$\bar Q\koszul$.
\item \label{item:boundary}
      The number~$b$ of boundary components of~$\surface_{\bar Q}$ can be computed as follows.
      There are two natural perfect matchings whose vertices are the blossom vertices of~$\bar Q$: one is obtained by joining the endpoints of each finite straight walk of~$\bar Q$, and the other is obtained similarly from~$\bar Q \koszul$.
      Let~$G$ be the superposition of these two perfect matchings.
      Then the number~$b$ of boundary components of~$\surface_{\bar Q}$ is the number of connected components of~$G$.
\item The genus of the surface~$\surface_{\bar Q}$ is
      \[
       g = \frac{|Q_1| - |Q_0| - b - p - p\dual + 2}{2},
      \]
      where~$b$ is the number of boundary components (see above for a way to compute~$b$) and~$p$ the number of punctures (\ie infinite straight walks in~$\bar Q$ and in~$\bar Q\koszul$).
      Indeed, $2g = 2-\chi(\surface_{\bar Q})$, where~$\chi(\surface_{\bar Q})$ is the Euler characteristic of~$\surface_{\bar Q}$.
      The dissection~$\dissection$ provides a cellular decomposition of~$\surface_{\bar Q}$ on which we fill the boundary components and the punctures with disks.  
      The number of faces of this cellular decomposition is~$b+p$, its number of edges is~$|Q_0|$ and its number of vertices is the number of finite straight walks in~$\bar Q$, which is equal to~$2|Q_0| - |Q_1|$.
      The above formula for~$g$ follows.
\end{enumerate}
\end{remark}

@@

\begin{example}[Cambrian paths]
\label{exm:CambrianPath}
Consider a \defn{Cambrian quiver}, that is any orientation of a line with no relations (\aka type~$A$ quiver).
Two such quivers are represented in \fref{fig:CambrianPathsQuivers}.
Note that we can choose to position the blossom vertices in such a way that for any arrow~$\beta$, the arrows~$\alpha$ and~$\gamma$ such that~$\alpha\beta \in I$ and~$\beta\gamma \in I$ are on the right of~$\beta$.

>>>>

\begin{remark}
\label{rem:propertiesSurface}
The following observations are useful for the computation of examples.
\begin{enumerate}[(i)]
\item The set~$V_{\bar Q}$ has one vertex for each straight walk in~$\bar Q$ (equivalently, for each maximal path in~$\bar Q$).
      Finite straight walks yield vertices on the boundary of~$\surface_{\bar Q}$, while infinite cyclic straight walks in~$\bar Q$ yield punctures~$\surface_{\bar Q}$ in~$V_{\bar Q}$.
      We denote by~$p$ the number of infinite cyclic straight walks in~$\bar Q$.
\item \label{item:arcs}
      The dissection~$\dissection_{\bar Q}$ has one edge for each vertex~$v \in \bar Q_0$, obtained by concatenation of the sides~$\Enrt{\alpha} = \Enrs{\beta}$ and~$\Enrt{\alpha'} = \Enrs{\beta'}$ where~$v = t(\alpha) = s(\beta) = t(\alpha') = s(\beta')$, $\alpha\beta \notin I$ and~$\alpha'\beta' \notin I$. We denote by~$\edgeof(v)$ the edge of~$\dissection_{\bar Q}$ corresponding to~$v$.
\item The dissection~$\dissection_{\bar Q}$ has one $\ell$-cell for each straight walk of length~$\ell$ in~$\bar Q\koszul$.
\item Similar statements hold dually for~${{V\dual}\!\!_{\bar Q}}$ and~${\dissection\dual\!\!_{\bar Q}}$, and the notations~$p\dual$ and~$\dualedgeof(v)$ are defined similarly.
\item The number of punctures of~$\surface_{\bar Q}$ is the number~$p + p\dual$ of infinite straight walks in~$\bar Q$ and in~$\bar Q\koszul$.
\item \label{item:boundary}
      The number~$b$ of boundary components of~$\surface_{\bar Q}$ can be computed as follows.
      There are two natural perfect matchings whose vertices are the blossom vertices of~$\bar Q$: one is obtained by joining the endpoints of each finite straight walk of~$\bar Q$, and the other is obtained similarly from~$\bar Q \koszul$.
      Let~$G$ be the superposition of these two perfect matchings.
      Then the number~$b$ of boundary components of~$\surface_{\bar Q}$ is the number of connected components of~$G$.
\item The genus of the surface~$\surface_{\bar Q}$ is
      \[
       g = \frac{|Q_1| - |Q_0| - b - p - p\dual + 2}{2},
      \]
      where~$b$ is the number of boundary components (see above for a way to compute~$b$) and~$p$ the number of punctures (\ie infinite straight walks in~$\bar Q$ and in~$\bar Q\koszul$).
      Indeed, ${2g = 2-\chi(\surface_{\bar Q})}$, where~$\chi(\surface_{\bar Q})$ is the Euler characteristic of~$\surface_{\bar Q}$.
      The dissection~$\dissection$ provides a cellular decomposition of~$\surface_{\bar Q}$ on which we fill the boundary components with disks.  
      The number of faces of this cellular decomposition is~$b + p\dual$, its number of edges is~$|Q_0|$ and its number of vertices is the number of straight walks in~$\bar Q$, which is equal to~$2|Q_0| - |Q_1| + p$.
      The above formula for~$g$ follows.
\end{enumerate}
\end{remark}

@@

\begin{example}
\label{exm:CambrianPath}
Consider a \defn{Cambrian quiver}, that is any orientation of a line with no relations (\aka type~$A$ quiver).
Two such quivers are represented in \fref{fig:CambrianPathsQuivers}.
Note that we can choose to position the blossom vertices in such a way that for any arrow~$\beta$, the arrows~$\alpha$ and~$\gamma$ such that~$\alpha\beta \in I$ and~$\beta\gamma \in I$ are on the right of~$\beta$.
